Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in the different poly shape power section

<!-- Figure \ref{fig:poly-example} shows one of the ten lineups where the visual test rejected the residual plot. The data plot, in position three, is very obvious, and different from the nulls. The visual test was correct to reject. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of polynomial to test for, and the recommendation is to generically use 4 (XXX REF). However, the "M" and the "Triple-U" shapes constructed from the Hermite polynomials use power up to 6 and 18 respectively. All 10 lineups rejected by the visual test but not the RESET were of types "M" or "Triple-U". If the RESET test had been applied using the power up to 6, all the decisions for these lineups would have been to reject. The recommendation of the polynomial power for the RESET should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.  -->

```{r poly-example, eval = FALSE, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test. The data plot at panel three exhibits strong non-linear pattern. It has effect size $= 3.89$. Interestingly, all the lineups in this group are higher order polynomials (M, Triple-U), and it is apparent that the default RESET test under-performs. The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials. The only way for it to be sensitive to these type of departure is to know the order prior to testing!"}
# plot poly-270 
VI_MODEL$plot_lineup(vi_lineup$poly_270$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

Shape of the non-linearity pattern is another factor contributed to the simulation model. Figure \ref{fig:poly-power-j}

```{r poly-power-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$."}

poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "polynomial") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(shape = factor(shape, levels = c("S", "U", "M", "Triple-U"))) %>%
  select(shape, log_effect_size, reject, conv_or_not) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = shape), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Vivid")) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Non-linearity shape")
```

>>>>

@@

Shape of the non-linearity pattern is another factor contributed to the non-linearity model. According to Figure \ref{fig:poly-power-j}, conventional tests are significantly less powerful in testing "Triple-U" shape compared to other three shapes, though it is still more sensitive than the visual test. In fact, among the lineups containing non-linearity patterns used in our experiments, around `r vi_survey %>% filter(!attention_check) %>% filter(!null_lineup) %>% filter(type == "polynomial") %>% group_by(unique_lineup_id) %>% summarise(across(everything(), first)) %>% count(a = p_value < 0.05, b = conventional_p_value < 0.05) %>% mutate(prop = n / sum(n) * 100) %>% filter(a, !b) %>% pull(prop) %>% round(digits = 1)`% (two our of 576) are rejected by the visual test, but accepted by the conventional test.

Figure \ref{fig:poly-example} shows one of the two lineups where the visual test rejects the residual plot. The data plot, in position 18, is very obvious, and different from the nulls. The visual test is correct to reject. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969). However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18. Both lineups rejected by the visual test but not the RESET are of "Triple-U". If the RESET test had been applied using the power up to six, all the decisions for these lineups would have been to reject. The recommendation of the polynomial power for the RESET should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.

We expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape. From Figure \ref{fig:poly-power-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape. Interestingly, the "M" shape is as easy to be detected as the "U" shape. The readability of the shape do not strictly follow the degree of the polynomials.

```{r poly-power-j, fig.height = 4, fig.cap = "Comparison of power of conventional tests and visual tests on lineups containing four different non-linearity shapes. "}

poly_conv_sim %>%
  filter(x_dist == "uniform") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "polynomial") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
  select(shape, log_effect_size, reject, conv_or_not) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = shape), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(4, "Vivid")) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Non-linearity shape")
```

```{r poly-example, fig.height = 7, fig.width = 7, fig.cap = "Example of a non-linearity that is rejected by the visual test but not by the RESET test. The data plot at panel 18 exhibits strong non-linear pattern. It has effect size $= 3.96$. Interestingly, all the lineups in this group are higher order polynomials (Triple-U), and it is apparent that the default RESET test under-performs. The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials. The only way for it to be sensitive to these type of departure is to know the order prior to testing!"}
# plot poly-275
VI_MODEL$plot_lineup(vi_lineup$poly_275$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

### 2023-01-23 paper_comparison/paper_comparison/paper_comparison.tex
fill in the different poly shape power section

Shape of the non-linearity pattern is another factor contributed to the
simulation model. Figure \ref{fig:poly-power-j}

@@

\caption{Power of conventional tests and visual tests on lineups produced by the non-linearity model with four different shapes controlled by the parameter $j$.}\label{fig:poly-power-j}
\end{figure}

>>>>

Shape of the non-linearity pattern is another factor contributed to the
non-linearity model. According to Figure \ref{fig:poly-power-j},
conventional tests are significantly less powerful in testing
``Triple-U'' shape compared to other three shapes, though it is still
more sensitive than the visual test. In fact, among the lineups
containing non-linearity patterns used in our experiments, around 0.3\%
(two our of 576) are rejected by the visual test, but accepted by the
conventional test.

Figure \ref{fig:poly-example} shows one of the two lineups where the
visual test rejects the residual plot. The data plot, in position 18, is
very obvious, and different from the nulls. The visual test is correct
to reject. To understand why this is, one needs to return to the way the
RESET test is applied. It requires a parameter indicating degree of
fitted values to test for, and the recommendation is to generically use
four (\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape
constructed from the Hermite polynomials use power up to 18. Both
lineups rejected by the visual test but not the RESET are of
``Triple-U''. If the RESET test had been applied using the power up to
six, all the decisions for these lineups would have been to reject. The
recommendation of the polynomial power for the RESET should be revised,
perhaps. This illustrates the sensitivity of conventional testing to the
parameters, and it also points to a limitation that one needs to know
the data structure in order to set the parameters for the test.

We expect the ``U'' shape will be the easiest one to be detected by
subjects followed by the ``S'', ``M'' and ``Triple-U'' shape. From
Figure \ref{fig:poly-power-j}, it can be observed that the power curves
are mostly aligned with our expectation, except for the ``M'' shape.
Interestingly, the ``M'' shape is as easy to be detected as the ``U''
shape. The readability of the shape do not strictly follow the degree of
the polynomials.

@@

\caption{Comparison of power of conventional tests and visual tests on lineups containing four different non-linearity shapes. }\label{fig:poly-power-j}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-example-1}

\caption{Example of a non-linearity that is rejected by the visual test but not by the RESET test. The data plot at panel 18 exhibits strong non-linear pattern. It has effect size $= 3.96$. Interestingly, all the lineups in this group are higher order polynomials (Triple-U), and it is apparent that the default RESET test under-performs. The RESET test using the higher order polynomial parameter could be used, but this would change it's performance for lower order polynomials. The only way for it to be sensitive to these type of departure is to know the order prior to testing!}\label{fig:poly-example}
\end{figure}
